---
title: "SAP - Project"
subtitle: "Analysis of UFC fights"
output: pdf_document
author: "Patrik Kukić, Filip Penzar, Željko Antunović, Noa Margeta"
date: "2023-01-06"
---

```{r,echo=FALSE,message=FALSE,warning=FALSE}
library (knitr)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

## Initial data analysis
```{r}
total_fight_data = read.csv('../total_fight_data.csv', sep=';')
dim(total_fight_data)

```

```{r}
fighter_details = read.csv('../fighter_details.csv', sep=',')
dim(fighter_details)
```
```{r}
all <- merge(total_fight_data, fighter_details, by.x='R_fighter', by.y='fighter_name', all.x = TRUE)
all <- merge(all, fighter_details, by.x='B_fighter', by.y='fighter_name',
              all.x = TRUE, suffixes = c('.r','.b'))
```
```{r}
dim(all)
```

## Task 1: Can we expect a fight to end by knockout depending on the difference in arm length between the fighters?

The initial step in solving this task was to convert the weight, height and reach of both fighters from the imperial system to the metric system. One conversion is shown here, the other 5 conversions were made in the same way.
We ignored all datapoints with NA values.

```{r}

# Conversion of inches to cm
all$Height_cm.b = sapply(strsplit(as.character(all$Height.b),"'|\""),
         function(x){30.48*as.numeric(x[1]) + 2.54*as.numeric(x[2])})

```


```{r, echo=FALSE}
all$Height_cm.r = sapply(strsplit(as.character(all$Height.r),"'|\""),
         function(x){30.48*as.numeric(x[1]) + 2.54*as.numeric(x[2])})

all$Reach_cm.b = sapply(strsplit(as.character(all$Reach.b),"'|\""),
         function(x){2.54*as.numeric(x[1])})

all$Reach_cm.r = sapply(strsplit(as.character(all$Reach.r),"'|\""),
         function(x){2.54*as.numeric(x[1])})
```


```{r, echo=FALSE}
# Convert lbs to kg
all$Weight_kg.b = sapply(strsplit(as.character(all$Weight.b)," "),
         function(x){0.453592*as.numeric(x[1])})

all$Weight_kg.r = sapply(strsplit(as.character(all$Weight.r)," "),
         function(x){0.453592*as.numeric(x[1])})
```


```{r}
# Moving lines that have NA reach
all_without_na_in_reach <- subset(all, !is.na(Reach_cm.b))
all_without_na_in_reach <- subset(all_without_na_in_reach, !is.na(Reach_cm.r))
```


```{r}
# Only fights that ended in a knockout
all_only_knockouts = subset(all_without_na_in_reach, all_without_na_in_reach$win_by=="KO/TKO")
```



```{r}
# Calculating the difference in the reach of winners and losers
d = c()
for (i in 1:nrow(all_only_knockouts)) {
   row = all_only_knockouts[i,]
   diff = row$Reach_cm.r - row$Reach_cm.b
   if (row$Winner == row$R_fighter) {
     d = append(d, diff)
   } else {
     d = append(d, -diff)
   }
}

summary(d)
```

```{r}
boxplot(d, ylab='Reach difference [cm]', main='Reach difference')
```

In order to be able to apply the t-test, it is first necessary to check the normality of the data distribution.

```{r}
hist(d, main = 'Winner and loser reach difference', xlab = 'd[cm]')
```
```{r}
qqnorm(d, pch=1, frame=FALSE, main = 'Reach difference')
qqline(d, col='blue', lwd=2)
```

From the histogram and Q-Q plot, we can conclude that the data is normally distributed, and we apply the t-test.

* H0: The difference in reach between winners and losers is zero.
* H1: Winners have a greater reach than losers.

```{r}
t.test(d, alternatives='greater', mu=0, conf.level=0.95)
```

With the significance level $\alpha=0.05$, we can reject the hypothesis H0 in favor of the hypothesis H1.

## Task 2: Does the duration of fights (s) differ between individual categories?

First, we calculated the overall duration of the fight from the record of the fight format and the duration of the last round.

```{r, echo=FALSE}
# Function to convert the format "3 Rnd (5-5-5)" to a vector of integers [5 5 5]
parse_format <- function(format){
   if (format == "No Time Limit"){
     return("No Time Limit")
   }
   format = substring(format, 1, nchar(format)-1)
   format = unlist(strsplit(format, "(", fixed=TRUE))
   format = unlist(strsplit(format[2], '-', fixed=TRUE))
   return(as.integer(format))
}

# Function to convert time from "m:s" format to seconds
convert_string_time_to_seconds <- function(time){
   time = unlist(strsplit(time, ":", fixed=TRUE))
   time = as.integer(time[1])*60 + as.integer((time[2]))
   return (time)
}
```

```{r}
# Calculation of the total duration of the fight
fight_length <- function(parsed_format, last_round, last_round_time){
   if (parsed_format[1] == "No Time Limit"){
     return (convert_string_time_to_seconds(last_round_time))
   }
   if (last_round == 1){
     return(convert_string_time_to_seconds(last_round_time))
   }
   total_time = 0
   for (i in 1:(last_round-1)){
     total_time = total_time + parsed_format[i]*60
   }
  
   total_time = total_time + convert_string_time_to_seconds(last_round_time)
   return(total_time)
}

# Based on the row calculation of the total duration of the fight
time_from_row <- function(row){
   parsed_format = parse_format(row$Format)
   last_round = row$last_round
   last_round_time = row$last_round_time
   return (fight_length(parsed_format, last_round, last_round_time))
}

# Calculation of the fight duration vector for each row of the table
dur = c()
for (i in 1:nrow(all)){
   dur = append(dur, time_from_row(all[i,]))
}

# Adding a column for the total duration of the fight in seconds
all$Fight_duration_s <- dur
```


```{r}
# Grouping by category (separated by gender)
men_classes = c('Light Heavyweight', 'Open Weight', 'Lightweight', 'Heavyweight',
                 'Featherweight', 'Bantamweight', 'Welterweight', 'Middleweight',
                 'Flyweight')
women_classes = c("Women's Bantamweight", "Women's Strawweight",
                   "Women's Featherweight", "Women's Flyweight")

# The function for string s returns TRUE if it contains one of the previously mentioned classes (men_classes, women_classes)
filter_not_in_classes <- function(s){
   for (w in women_classes){
     if (grepl(w, s)) {
       return(TRUE)
     }
   }
   for (m in men_classes) {
     if (grepl(m, s)) {
       return(TRUE)
     }
   }
   return(FALSE)
}

# Function for string s returns the category from men_classes or women_classes it contains
check_which_class <- function(s){
   for (w in women_classes){
     if (grepl(w, s)) {
       return(w)
     }
   }
   for (m in men_classes) {
     if (grepl(m, s)) {
       return(m)
     }
   }
}

# All types of fights that we do not know how to group into categories by weight and gender
ignore_fight_types = c()
categories = unique(all$Fight_type)
for (category in categories) {
   if (!filter_not_in_classes(category)) {
     ignore_fight_types = append(ignore_fight_types, category)
   }
}

ignore_fight_types
```

Certain categories do not contain information about gender and weight, and therefore we do not take them into account during further analysis.

```{r}
# From the entire data set, we move the fights whose fight_type is inside the ignore_fight_types vector
all_without_unknown_weight_classes = subset(all, !(Fight_type %in% ignore_fight_types))
```

The assumptions of the parametric ANOVA method are:

* independence of individual data in the samples
* normal data distribution
* homogeneity of variances among populations

1) We assume independence of the data in the samples, because the fights are mutually independent.
2) We continue with testing the normality of data distribution. We use the Lilliefors normality test.
* H0: The data belong to a normal distribution.
* H1: The data does not belong to a normal distribution.
3) If the data distribution is not normal, there is no point in checking for homoscedasticity. In the second case, we have to check homoscedasticity with Bartlett's test.

```{r, echo=FALSE}
# Adding a category column
weight_class = c()
for (i in 1:nrow(all_without_unknown_weight_classes)) {
   weight_class = append(weight_class, check_which_class(all_without_unknown_weight_classes[i,]$Fight_type))
}
all_without_unknown_weight_classes$weight_class = weight_class
```

```{r, echo=FALSE}
require(nortest)
```
```{r}
lillie.test(all_without_unknown_weight_classes$Fight_duration_s[weight_class == "Open Weight"])
```

Due to the very small $p$ value, we reject H0 in favor of H1 and conclude that the data are not normally distributed. That's why we have to use the non-parametric version of the ANOVA test, the Kruskal-Wallis $\chi^2$-test. Therefore, we do not test for homogeneity of variances across categories.

```{r}
# weight_classes = c(men_classes, women_classes)
hist(all_without_unknown_weight_classes$Fight_duration_s[all_without_unknown_weight_classes$weight_class == "Open Weight"],
      xlab="Fight duration [s]", main="Open Weight")
hist(all_without_unknown_weight_classes$Fight_duration_s[all_without_unknown_weight_classes$weight_class == "Heavyweight"],
      xlab="Fight duration [s]", main="Heavyweight")
hist(all_without_unknown_weight_classes$Fight_duration_s[all_without_unknown_weight_classes$weight_class == "Women's Flyweight"],
      xlab="Fight duration [s]", main="Women's Flyweight")

```
From the histograms shown, we can see that the fight times are consistent with the fight formats (most fights end in the 15th minute because they are 5+5+5 minutes).


In order to perform the Kruskal-Wallis test, we must have a minimum of 5 observations in each of the categories, which we can confirm from the following table:

```{r}
table(all_without_unknown_weight_classes$weight_class)
```

We put forward hypotheses:

* H0: The duration of fights does not differ between categories.
* H1: Fight duration differs between at least two categories.
 
```{r}
kruskal.test(Fight_duration_s~weight_class, data=all_without_unknown_weight_classes)
```

Due to the small $p$-value, we reject H0 in favor of H1 and conclude that the duration of fights is statistically significantly different between at least two weight categories.


## Task 3: Do title fights last longer (in rounds) than other fights in the competition?

```{r, echo=FALSE}
# Function that returns TRUE if "title" is in the string s
is_title_bout <- function(s) {
   return(grepl('title', s, ignore.case=TRUE))
}

# A vector containing TRUE if the fight in the row is a title fight
title_bout_vector = c()
for (i in (1:nrow(all))) {
   title_bout_vector = append(title_bout_vector, is_title_bout(all[i,]$Fight_type))
}
# Vector of number of rounds for title fights
title_bouts_last_round = all[title_bout_vector,]$last_round

# Vector of round counts for non-title fights
non_title_bouts_last_round = all[!title_bout_vector,]$last_round
```

```{r}
hist(title_bouts_last_round, breaks=seq(min(title_bouts_last_round)-0.5, max(title_bouts_last_round)+0.5, by=1),
      main="Freq. of number of rounds for title bouts", xlab="No. of rounds")
hist(non_title_bouts_last_round, breaks=seq(min(non_title_bouts_last_round)-0.5, max(non_title_bouts_last_round)+0.5, by=1),
      main="Freq. of number of rounds for non title bouts", xlab="No. of rounds")
```

```{r}
lillie.test(non_title_bouts_last_round)
lillie.test(title_bouts_last_round)
```


From the histogram and the Lilliefors test, we see that the data are not normally distributed, so we apply the non-parametric version of the t-test, the Wilcoxon signed rank test.
We hypothesize:
- H0: Title fights do not last longer (in rounds) than other fights in the competition.
- H1: Title fights last longer (in rounds) than other fights in the competition.

```{r}
wilcox.test(title_bouts_last_round, non_title_bouts_last_round, alternatives='greater', conf.level=0.9)
```

We chose the level of significance $\alpha=0.1$ because we want greater robustness of the test.
Due to the calculated $p$-value, we reject H0 in favor of H1 and conclude that title fights last longer (in rounds) than other fights in the competition.

## Additional task 1. - Do shorter fighters win more often via submission?

We have added a gender attribute to all fighters, which we have determined through the names of the fight categories in which that fighter fought.

```{r, echo=FALSE}
# Function to determine the gender of fighters (via category)
male_or_female = c()
for (i in 1:nrow(all_without_unknown_weight_classes)) {
   if (all_without_unknown_weight_classes[i,]$weight_class %in% men_classes) {
     male_or_female = append(male_or_female, 'male')
   } else {
     male_or_female = append(male_or_female, 'female')
   }
}

# Adding gender column
all_without_unknown_weight_classes$male_or_female = male_or_female
```

```{r, echo=FALSE}
# Adding gender to fighter_details table
total_gender_vector = c()
for (i in 1:nrow(fighter_details)) {
   gender_vector = all_without_unknown_weight_classes[fighter_details[i,]$fighter_name == all_without_unknown_weight_classes$B_fighter
                                                 | fighter_details[i,]$fighter_name == all_without_unknown_weight_classes$R_fighter,]$male_or_female
   if (length(gender_vector) == 0) {
     total_gender_vector = append(total_gender_vector, NA)
   } else {
     total_gender_vector = append(total_gender_vector, gender_vector[1])
   }
}

fighter_details$gender = total_gender_vector
```

Then we converted the sum of all the fighters' heights from the imperial system of measurement units to the metric system.

```{r, echo=FALSE}
# Conversion of inches to cm for fighter height
fighter_details$Height_cm = sapply(strsplit(as.character(fighter_details$Height),"'|\""),
         function(x){30.48*as.numeric(x[1]) + 2.54*as.numeric(x[2])})
```

```{r}
female_heights = subset(fighter_details, gender == 'female')$Height_cm
male_heights = subset(fighter_details, gender == 'male')$Height_cm
hist(female_heights, main="Female heights", xlab="Height[cm]")
hist(male_heights, main="Male heights", xlab="Height[cm]")

```

Based on the median of all male and female heights, we divided the fighters into shorter and taller fighters, with regard to gender.

```{r}
# Determining the median height for the male and female population
male_median_height = median(male_heights, na.rm=TRUE)
female_median_height = median(female_heights, na.rm=TRUE)

# Determination of height category by gender (short for heights below the median, tall for heights above the median)
height_category = c()
for (i in 1:nrow(fighter_details)){
   if (is.na(fighter_details[i,]$Height_cm) | is.na(fighter_details[i,]$gender)) {
       height_category = append(height_category, NA)
       next
   }
   else {
     if (fighter_details[i,]$gender == "male"){
       if (fighter_details[i,]$Height_cm >= male_median_height){
         height_category = append(height_category, "tall")
       }
       else{
         height_category = append(height_category, "short")
       }
     }else{
       if (fighter_details[i,]$Height_cm >= female_median_height){
         height_category = append(height_category, "tall")
       }
       else{
         height_category = append(height_category, "short")
       }
     }
   }
}

# Adding a height category column
fighter_details$height_category = height_category
```

For each fighter, we determined the percentage of his victories through submission of the opposing fighter. If the fighter did not have a single victory, we marked the percentage of victories by submission of the opponent with NA.

```{r, echo=FALSE}
# Determining the percentage through submission for each fighter
win_rate_by_submission = c()
for (i in (1:nrow(fighter_details))) {
   fn = fighter_details[i,]$fighter_name
   wins = subset(all_without_unknown_weight_classes, Winner==fn)
   n = nrow(wins)
   if (n == 0){
     win_rate = NA
   } else {
     won_by_submission = subset(wins, win_by=='Submission')
     s = nrow(won_by_submission)
     win_rate = s / n
   }
   win_rate_by_submission = append(win_rate_by_submission, win_rate)
}

# Adding a submission win percentage column
fighter_details$win_rate_by_submission = win_rate_by_submission
```

```{r}
# Vector of submission win percentages for low fighters
short_winners = subset(fighter_details, height_category=='short' &
                          !is.na(win_rate_by_submission))$win_rate_by_submission
# Vectors of submission win percentages for tall fighters
tall_winners = subset(fighter_details, height_category=='tall' &
                         !is.na(win_rate_by_submission))$win_rate_by_submission

hist(short_winners, main="Freq. of short fighters submission win rate", xlab="Win rate by submission")
hist(tall_winners, main="Freq. of tall fighters submission win rate", xlab="Win rate by submission")
```
We propose the following hypotheses:

* H0: Percentages of submission wins are the same for tall and short fighters.
* H1: The percentage of submission wins is lower for tall fighters.

We set the significance level $\alpha$ to 0.1 because we want to be less sensitive to not rejecting H0.

```{r}
wilcox.test(tall_winners, short_winners, alternatives='less', conf.level=0.9)
```

At the level of significance $\alpha=0.1$ and the $p$ value obtained from the Wilcoxon rank sum test, we conclude that we cannot reject H0 in favor of H1 (we cannot reject the hypothesis that the percentages of submission wins are the same for high and low fighters).

```{r, echo=FALSE}
# short_winners_without_outliers = short_winners[short_winners != 0 & short_winners != 1]
# tall_winners_without_outliers = tall_winners[tall_winners != 0 & tall_winners != 1]
# hist(short_winners_without_outliers)
# hist(tall_winners_without_outliers)
# wilcox.test(tall_winners_without_outliers, short_winners_without_outliers, alternatives='less', conf.level=0.9)
```

## Additional task 2. - Do men's fights end more often with a knockout?

As with the previous task, we first determined each fighter's knockout percentage. For a fighter without a win, we recorded the percentage of wins by knockout with NA.

```{r, echo=FALSE}
# Determining the percentage of victory by knockout for each fighter
win_rate_by_ko = c()
for (i in (1:nrow(fighter_details))) {
   fn = fighter_details[i,]$fighter_name
   wins = subset(all_without_unknown_weight_classes, Winner==fn)
   n = nrow(wins)
   if (n == 0){
     win_rate = NA
   } else {
     wins_by_ko = subset(wins, win_by=='KO/TKO')
     s = nrow(wins_by_ko)
     win_rate = s / n
   }
   win_rate_by_ko = append(win_rate_by_ko, win_rate)
}

# Adding a column with the percentage of victory by knockout
fighter_details$win_rate_by_ko = win_rate_by_ko
```


```{r}
female_ko_winners = subset(fighter_details, gender=='female' & !is.na(win_rate_by_ko))$win_rate_by_ko
male_ko_winners = subset(fighter_details, gender=='male' & !is.na(win_rate_by_ko))$win_rate_by_ko

hist(female_ko_winners, main="Freq. of female fighters knockout win rate", xlab="Win rate by knockout")
hist(male_ko_winners, main="Freq. of male fighters knockout win rate", xlab="Win rate by knockout")
```
We hypothesize:

* H0: The percentages of victories by knockout are equal for men and women.
* H1: The percentage of victories by knockout is higher for men.

We set the significance level $\alpha$ to 0.1, as in the previous tests.

```{r}
wilcox.test(male_ko_winners, female_ko_winners, alternatives='greater', conf.level=0.9)
```
At the significance level $\alpha=0.1$, we can reject H0 in favor of H1 (the percentage of victories by knockout is higher for men).

```{r, echo=FALSE}
# male_ko_winners_without_outliers = male_ko_winners[male_ko_winners != 0 & male_ko_winners != 1]
# female_ko_winners_without_outliers = female_ko_winners[female_ko_winners != 0 & female_ko_winners != 1]
#
# hist(male_ko_winners_without_outliers)
# hist(female_ko_winners_without_outliers)
#
# wilcox.test(male_ko_winners_without_outliers, female_ko_winners_without_outliers, alternatives='less', conf.level = 0.9)
# wilcox.test(male_ko_winners_without_outliers, female_ko_winners_without_outliers, alternative='greater', conf.level = 0.9)
```

## Additional task 3. - Does the number of victories and victories by knockout differ depending on the attitude of the fighter (stance)?

```{r}
# Determination of the number of victories and the number of victories by knockout for fighters
total_wins = c()
total_wins_by_ko = c()
for (i in (1:nrow(fighter_details))) {
   fn = fighter_details[i,]$fighter_name
   wins = subset(all, Winner==fn)
   wins_by_ko = subset(wins, win_by=='KO/TKO')
   total_wins = append(total_wins, nrow(wins))
   total_wins_by_ko = append(total_wins_by_ko, nrow(wins_by_ko))
}

# Adding total wins, total wins by knockout and total wins without knockout columns
fighter_details$total_wins = total_wins
fighter_details$total_wins_by_ko = total_wins_by_ko
fighter_details$total_wins_without_ko = total_wins - total_wins_by_ko

table(fighter_details$Stance)
```


We ignore fighters with an unknown attitude. We also ignore fighters with "Open Stance" and "Sideways" due to low frequency.
If he is a fighter of the "Orthodox" position, then he is right-handed. If the stance is "Southpaw", then he is left-handed. If it's "Switch", then it's ambidextrous.

```{r, echo=FALSE}
fighter_details_known_stance = subset(fighter_details, Stance != '' & Stance!='Open Stance' & Stance != 'Sideways')

library(dplyr)
library(data.table)
stance_table <- fighter_details_known_stance %>%
   group_by(Stance) %>%
   summarize(total_wins_by_ko = sum(total_wins_by_ko), total_wins_without_ko = sum(total_wins_without_ko)) %>%
   setDT()
```

```{r}
stance_table
```

```{r}
# We have to remove Stance because in the table it is represented as a dependent variable, but in fact it is an independent variable
stance_table = select(stance_table, -Stance)
```

The expected frequencies are greater than 5 in each cell of the table. Therefore, we can apply the homogeneity test.
We hypothesize:

* H0: The percentage of victories by knockout is the same for each category of fighters according to stance (left-handed, right-handed and ambidextrous).
* H1: The percentage of victories by knockout is not equal for at least two of the categories of fighters according to stance (left-handed, right-handed and ambidextrous).

No conf_level argument is available for chisq.test, so we do not set any significance level as a test argument. However, we choose a significance level of $\alpha=0.05$.

```{r}
chisq.test(stance_table, correct=FALSE)
```

At the selected significance level, we can reject H0 in favor of H1 (the proportion of wins by KO and wins by other means is not the same for all Stance categories).
From the *stance_table* we can guess that fighters who are ambidextrous have a higher proportion of victories by KO.

## Task 4: Can we predict the winner from the given characteristics?

For each fight, we calculated the age of both fighters (Red and Blue) on the day of the fight.

```{r, echo=FALSE}
# Calculation of the current age of both fighters on the date of the fight
library (lubridate)
current_age_blue = c()
current_age_red = c()
fight_date = mdy(all_without_unknown_weight_classes[1,]$date) #converting string to date
red_date_of_birth = mdy(all_without_unknown_weight_classes[1,]$DOB.r)
age = fight_date - red_date_of_birth
for (i in (1:nrow(all_without_unknown_weight_classes))){
   fight_date = mdy(all_without_unknown_weight_classes[i,]$date)
   red_date_of_birth = mdy(all_without_unknown_weight_classes[i,]$DOB.r)
   blue_date_of_birth = mdy(all_without_unknown_weight_classes[i,]$DOB.b)
   red_age = as.numeric(floor((fight_date - red_date_of_birth)/365))
   blue_age = as.numeric(floor((fight_date - blue_date_of_birth)/365))
   current_age_red = append(current_age_red, red_age)
   current_age_blue = append(current_age_blue, blue_age)
}

# Adding current age column
all_without_unknown_weight_classes$red_age = current_age_red
all_without_unknown_weight_classes$blue_age = current_age_blue
```

Certain columns within the table of all fights are in the form "*x* of *y*" because they tell how many punches the fighter saved, received, etc.
For a subset of those columns, we only considered the first number *x*, because it gives us information about the blows exchanged during the fight.
Another subset of those columns describes the fighter's overall accuracy, and for that subset of columns we calculated the *x*/*y* ratio (percentage).

```{r, echo=FALSE}
# The only fights we take into account are those with a known weight category
all_for_logreg = data.frame(all_without_unknown_weight_classes)

# Funkcija za pretvorbu stringa s oblika "x of y" u numeric x
number_before_of <- function(s) {
  first_elem = strsplit(s, " of ", fixed=TRUE)[[1]][1]
  return(as.numeric(first_elem))
}
```

```{r, echo=FALSE}
# Pretvorba stupaca iz oblika "x of y" u "x"
all_for_logreg$r_sig_str = sapply(all_for_logreg$R_SIG_STR., number_before_of)
all_for_logreg$b_sig_str = sapply(all_for_logreg$B_SIG_STR., number_before_of)
all_for_logreg$r_total_str = sapply(all_for_logreg$R_TOTAL_STR., number_before_of)
all_for_logreg$b_total_str = sapply(all_for_logreg$B_TOTAL_STR., number_before_of)

to_transform_colnames <- c("R_TD", "B_TD", "R_HEAD", "B_HEAD", "R_BODY", "B_BODY", "R_LEG", "B_LEG", "R_DISTANCE", "B_DISTANCE", "R_CLINCH", "B_CLINCH", "R_GROUND", "B_GROUND")
for (column in to_transform_colnames) {
  new_column_name = tolower(column)
  all_for_logreg[,new_column_name] = sapply(all_for_logreg[,column], number_before_of)
}

# Dodavanje indikatorske varijable o pobjedi crvenog borca
all_for_logreg$red_is_winner = as.numeric(all_for_logreg$R_fighter == all_for_logreg$Winner)
```

```{r, echo=FALSE}
# Funkcija za pretvaranje stringa s oblika "x%" u numeric "x/100"
number_before_percent <- function(s) {
  return (as.numeric(sub("%","",s))/100)
}

# Pretvorba stupaca iz oblika "x%" u numeric "x/100"
to_transform_colnames = c("Str_Def.r", "Str_Def.b", "Str_Acc.r", "TD_Acc.r", "TD_Def.r", "Str_Def.b", "Str_Acc.b", "TD_Acc.b", "TD_Def.b")
for (column in to_transform_colnames) {
  new_column_name = tolower(column)
  all_for_logreg[,new_column_name] = suppressWarnings(sapply(all_for_logreg[,column], number_before_percent))
}

# Uzimamo u obzir samo borce koji imaju stav: "Switch", "Orthodox" i "Southpaw"
all_for_logreg = subset(all_for_logreg, Stance.b=="Switch" | Stance.b == "Orthodox" | Stance.b=="Southpaw")
all_for_logreg = subset(all_for_logreg, Stance.r=="Switch" | Stance.r == "Orthodox" | Stance.r=="Southpaw")

# Pretvaranje logičkog vektora u numerički vektor
all_for_logreg$is_b_southpaw = as.numeric(all_for_logreg$Stance.b == "Southpaw")
all_for_logreg$is_b_orthodox = as.numeric(all_for_logreg$Stance.b == "Orthodox")
all_for_logreg$is_b_switch = as.numeric(all_for_logreg$Stance.b == "Switch")
all_for_logreg$is_r_southpaw = as.numeric(all_for_logreg$Stance.r == "Southpaw")
all_for_logreg$is_r_orthodox = as.numeric(all_for_logreg$Stance.r == "Orthodox")
all_for_logreg$is_r_switch = as.numeric(all_for_logreg$Stance.r == "Switch")
```

Nakon toga smo odredili regresorske varijable. Zavisna varijabla je indikatorska varijabla u obliku vektora (označava pobjedu crvenog borca).

```{r}
# Odabrane regresorske varijable i zavisna varijabla
selected_columns = c("R_KD", "B_KD", "R_SUB_ATT", "B_SUB_ATT", "R_REV", "B_REV", "TD_Avg.r", "SLpM.r", "SApM.r", "Sub_Avg.r", "TD_Avg.b", "SLpM.b", "SApM.b", "Sub_Avg.b", "Height_cm.b", "Height_cm.r", "Reach_cm.b", "Reach_cm.r", "Weight_kg.b", "Weight_kg.r", "red_age", "blue_age", "r_sig_str", "b_sig_str", "r_total_str", "b_total_str", "r_td", "b_td", "r_head", "b_head", "r_body", "b_body", "r_leg", "b_leg", "r_distance", "b_distance", "r_clinch", "b_clinch", "r_ground", "b_ground", "str_def.r", "str_acc.r", "td_acc.r", "td_def.r", "str_def.b", "str_acc.b", "td_acc.b", "td_def.b", "red_is_winner", "is_b_southpaw", "is_b_orthodox", "is_r_southpaw", "is_r_orthodox")
variables = selected_columns[selected_columns != "red_is_winner"]
```


```{r}
library(tidyr)
# Iz seta podataka uzimamo samo odabrane regresorske varijable i zavisnu varijablu
logreg_data = subset(all_for_logreg, select=selected_columns)
# Uzimamo samo retke koji nemaju NA vrijednosti unutar odabranih varijabli
logreg_data = logreg_data %>% drop_na()
```

Koristimo model logističke regresije jer je zavisna varijabla indikatorska.
```{r}
require(caret)

# b je formula varijabla_1 + varijabla_2 + ..., pri čemu je varijabla_i unutar 
# skupa odabranih regresorskih varijabli
b <- paste(variables, collapse=' + ')
logreg_mdl = glm(as.formula(paste("red_is_winner ~ ", b)), data=logreg_data, family=binomial())
summary(logreg_mdl)
```
Iz ispisa uočavamo da su neki od regresora međusobno zavisni (NA vrijednosti). U ispisu su označeni statistički signifikantni regresori. 

Na tri različita načina evaluirat ćemo kvalitetu dobivenog modela.

Računamo $R^2$ koji govori o tome koliko je procjenjeni model blizu ili daleko od nul-modela (što je $R^2$ bliži 1, to je model bolji).

```{r}
# Računanje Rsq
Rsq = 1 - logreg_mdl$deviance/logreg_mdl$null.deviance
Rsq
```

Izrađujemo matricu zabune.
```{r}
# Izrada confusion matrix-a
yhat <- logreg_mdl$fitted.values >= 0.5
tab <- table(logreg_data$red_is_winner, yhat)

tab
```

Iz matrice zabune možemo zaključiti da model dobro predviđa ishod borbe (borbe u kojima crveni borac nije pobjednik su označene kao takve, i obrnuto).

```{r}
accuracy = sum(diag(tab))/sum(tab)
precision = tab[2,2]/sum(tab[,2])
recall = tab[2,2]/sum(tab[2,])
specificity = tab[1,1]/sum(tab[,1])

accuracy
precision
recall
specificity
```
Due to the high values of the calculated variables (accuracy, precision, response and specificity), we conclude that the model is of high quality.

## Model without linearly dependent and insignificant regressors

```{r}
# Dropping non-significant variables
significant_variables = c("R_KD", "B_KD", "R_SUB_ATT", "B_SUB_ATT", "R_REV", "B_REV",
                           "TD_Avg.r", "red_age", "r_sig_str", "b_sig_str", "r_total_str",
                           "r_td", "b_td", "r_head", "b_head", "r_distance", "b_distance",
                           "r_clinch", "b_clinch", "td_acc.b")

b <- paste(significant_variables, collapse=' + ')
logreg_mdl_reduced = glm(as.formula(paste("red_is_winner ~ ", b)), data=logreg_data, family=binomial())
summary(logreg_mdl_reduced)
```
As for the previous model, we calculate the same quality measures ($R^2$, accuracy, precision, response and specificity).
```{r}
Rsq = 1 - logreg_mdl_reduced$deviance/logreg_mdl_reduced$null.deviance
Rsq
```

```{r}
yhat <- logreg_mdl_reduced$fitted.values >= 0.5
tab <- table(logreg_data$red_is_winner, yhat)

tab

accuracy = sum(diag(tab))/sum(tab)
precision = tab[2,2]/sum(tab[,2])
recall = tab[2,2]/sum(tab[2,])
specificity = tab[1,1]/sum(tab[,1])

accuracy
precision
recall
specificity
```

## Comparison of original and reduced model

We will use ANOVA to compare the models.
We hypothesize:

* H0: The models are of equal quality
* H1: The original model is better than the reduced one

```{r}
# Comparison of two models
anova(logreg_mdl, logreg_mdl_reduced, test='LRT')
```

With the significance level $\alpha=0.05$, we conclude that we can reject H0 in favor of H1 (the original model is better than the reduced one).

## Model with a priori data

An interesting question arises whether we can determine the winner based only on the features available before the fight (previous statistics of each fighter).

We select only the variables available before the fight for each fighter, and use them as regressors in the new logistic model.
```{r}
fighter_details_variables = c("TD_Avg.r", "SLpM.r", "SApM.r", "Sub_Avg.r", "TD_Avg.b",
                               "SLpM.b", "SApM.b", "Sub_Avg.b", "Height_cm.b",
                               "Height_cm.r", "Reach_cm.b", "Reach_cm.r", "Weight_kg.b",
                               "Weight_kg.r", "red_age", "blue_age", "str_def.r", "str_acc.r",
                               "td_acc.r", "td_def.r", "str_def.b", "str_acc.b", "td_acc.b",
                               "td_def.b", "red_is_winner", "is_b_southpaw", "is_b_orthodox",
                               "is_r_southpaw", "is_r_orthodox")
logreg_fighters_data = subset(logreg_data, select = fighter_details_variables)
fighter_details_variables = fighter_details_variables[fighter_details_variables != "red_is_winner"]

b <- paste(fighter_details_variables, collapse=' + ')
logreg_mdl_fighter_details = glm(as.formula(paste("red_is_winner ~ ", b)),
                                  data=logreg_fighters_data, family=binomial())
summary(logreg_mdl_fighter_details)
```

We calculate measures of model quality.
```{r}
# Calculation of Rsq
Rsq = 1 - logreg_mdl_fighter_details$deviance/logreg_mdl_fighter_details$null.deviance
Rsq
```
```{r}
yhat <- logreg_mdl_fighter_details$fitted.values >= 0.5
tab <- table(logreg_fighters_data$red_is_winner, yhat)

tab
```
```{r}
accuracy = sum(diag(tab))/sum(tab)
precision = tab[2,2]/sum(tab[,2])
recall = tab[2,2]/sum(tab[2,])
specificity = tab[1,1]/sum(tab[,1])

accuracy
precision
recall
specificity
```
From the calculated quality measures, we infer that the model is worse than the previous ones, but also that it is better than ordinary guessing.